---
title: "Untitled"
author: "Zachary M. Smith"
date: "4/4/2020"
output: html_document
---

# Methods

## stayCALM Package Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r install-github, eval=FALSE}
# install.packages("devtools")
devtools::install_github("BWAM/stayCALM")
```

You must have `rtools` installed on your machine for the stayCALM to be built upon installation from GitHub. `rtools` is __NOT__ an R package. On Windows machines, `rtools` can be installed from the following CRAN repository: https://cran.r-project.org/bin/windows/Rtools/.


## Prepare the R Environment

Load the necessary R packages into the global environment, including the __stayCALM__ package.
```{r load-packages, message=FALSE}
library(tidyverse)
library(stayCALM)
```

## Preprocess Data

### Water Quality Standards

Data was preprocessed to resemble the expected output from the athorotative databases that will become available as part of the Data Modernization effort [Data Modernization].
```{r}
data(nysdec_wqs)
```

- __fresh__ - fresh surface waters
    + D > C > B > A > A-Special > AA > AA-Special
    + D > C > C(T) > C(TS) > B > B(T) > B(TS) > A > A(T) > A(TS) > A-Special  > A-Special(T) > A-Special(TS) > AA > AA(T) > AA(TS) > AA-Special > AA-Special(T) > AA-Special(TS)
- __saline__ - saline surface waters
    + SD > I > SC > SB > SA
```{r}
wqs.df <- purrr::map_df(c("fresh", "saline", "saline_i"), function(type.i) {
  class.vec <- switch(type.i,
                      "fresh" = c("D", class_variants(.class_vec = c("C", "B",
                                                                     "A", "A-Special",
                                                                     "AA", "AA-Special"),
                                                      .variant_vec = c("", "(T)", "(TS)"))),
                      "saline" = c("SD", "I", "SC", "SB", "SA"))
                      #"saline_i" = "I")
  
  wqs_sub.df <- nysdec_wqs[nysdec_wqs$class_type == type.i, ]
  
  
  final.df <- class_inheritance(.data = wqs_sub.df,
                                parameter,
                                .class_col = class,
                                .levels_vec = class.vec) %>% 
    dplyr::mutate(class = as.character(class))
    
})
```

```{r, echo=FALSE}
standard_dt(wqs.df)
```

### WI/PWL

Data was preprocessed to resemble the expected output from the athorotative databases that will become available as part of the Data Modernization effort [Data Modernization].
```{r}
data("wipwl.df")
```

```{r}
wqs_wipwl.df <- wipwl.df %>%
  select(seg_id, class, tmdl_id, mgmtstat) %>% 
  distinct() %>%
  left_join(wqs.df, by = "class")
```

### Stream Data

Data was preprocessed to resemble the expected output from the athorotative databases that will become available as part of the Data Modernization effort [Data Modernization].
```{r}
data(stream_chem.df) 
```


```{r}
chem.df <- stream_chem.df %>% 
  rename_all(tolower) %>% 
  rename(seg_id = sh_pwl_id,
         parameter = chemical_name) %>% 
  inner_join(wqs_wipwl.df, by = c("seg_id", "parameter"))
```


```{r}
chem_final.df <- chem.df %>% 
  mutate(result_numeric = if_else(is.na(result_numeric) & !is.na(validation_level),
                                  reporting_detection_limit,
                                  result_numeric)) %>% 
    filter(!validation_level %in% "Not Reported",
         !is.na(result_numeric))
```

## Assess

Thresholds represented by `NA` will throw an error. The majority of the `NA`s in the threshold column now need to be calculated-- a task I have not standardized yet.
```{r}
chem_final.df <- chem_final.df %>%
  filter(!is.na(threshold))
```

```{r}
assessed.df <- chem_final.df %>%
  # Drop this filter after additional prep steps have been added
  filter(is.na(formula) & statistic %in% "none" & block %in% "single") %>% 
  date_standard_cols(.date_col = sample_date) %>% 
  assessment_id(seg_id,
                lab_anl_method_name,
                cas_rn,
                parameter) %>%
  mutate(
    within_period = assessment_period(.df = .,
                      .date_col = date,
                    .n_years_ago = 10),
    min_years_samples = all(assessment_min_years(.df = .,
                                                .date_col = date,
                                                .min_years = 2,
                                                assessment_id),
                           assessment_min_counts(.df = ., 
                                                 .min_samples = 8,
                                                 assessment_id)),
    attaining_wqs = wqs_attaining(result_numeric,
                                  direction,
                                  threshold,
                                  na.rm = TRUE),
    attaining_75 = wqs_attaining(result_numeric,
                                 direction,
                                 threshold,
                                 na.rm = TRUE),
    min_violations_year = TRUE,
    tmdl = !is.na(tmdl_id),
    ltco_rest_plan = mgmtstat %in% "Strategy Implementation Scheduled or Underway",
    pollutant = TRUE 
    )
```



```{r}
test.df <- assessed.df %>% 
  mutate(
    assessment = case_when(
      !within_period ~ "Manual Review Required: Manual Review Required: Manual Review Required",
      !min_years_samples & attaining_wqs ~ "IR3: Impaired: Unconfirmed",
      !min_years_samples & !attaining_wqs & attaining_75  ~ "IR3: Fully Supported: Unconfirmed",
      !min_years_samples & !attaining_wqs & !attaining_75  ~ "IR3: Stressed: Unconfirmed",
      min_years_samples & attaining_wqs & attaining_75  ~ "IR1: Fully Supported: Confirmed",
      min_years_samples & attaining_wqs & !attaining_75  ~ "IR1: Stressed: Confirmed",
      min_years_samples & !attaining_wqs & !min_violations_year  ~ "IR3: Impaired: Unconfirmed",
      min_years_samples & !attaining_wqs & min_violations_year & tmdl  ~ "IR4a: Impaired: Confirmed",
      min_years_samples & !attaining_wqs &
          min_violations_year & !tmdl & ltco_rest_plan  ~ "IR4b: Impaired: Unconfirmed",
      min_years_samples & !attaining_wqs &
          min_violations_year & !tmdl & !ltco_rest_plan & pollutant  ~ "IR4c: Impaired: Confirmed",
      min_years_samples & !attaining_wqs &
          min_violations_year & !tmdl & !ltco_rest_plan & !pollutant  ~ "IR5: Impaired: Confirmed",
      TRUE ~ "ERROR"
    )
  )

# table(test.df$assessment)
```